---
title: "Efficient green energy sources report"
author: "Alejandro Barajas, Gary Layman, Trevor Olmo, Tyler Steppat, Shihan Embuldeniya"
date: today
format: pdf
editor: source
---

# R and Python libary set up

```{r}
#| label: r setup
#| include: false
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, results = "hold")
library(tidyverse)
library(reticulate)
library(eia)
eia_set_key(Sys.getenv("EIA_KEY"))

# setup python environment
use_virtualenv("./.venv")
```

```{python}
#| label: python setup
#| include: false
import pandas as pd
from glob import glob
import os
```

- **Overall goal**
The goal is to perform a comparative analysis of energy sources to determine which produces the most energy per dollar while minimizing carbon emissions.

## Data Preparation
```{r}
#| label: Fetch EIA data
#### Commented this out because we no longer need it
# energy <- eia_data(
# dir = "electricity/electric-power-operational-data",
#  data = c("ash-content", "consumption-for-eg", "consumption-for-eg-btu", "consumption-uto", "consumption-uto-btu", "cost", "cost-per-btu", "generation", "heat-content", "receipts", "receipts-btu", "stocks", "sulfur-content", "total-consumption", "total-consumption-btu"),
#  freq = "quarterly",
#  start = "2020",
#  end = "2025"
#)
# All columns are <char> and will need to be cleaned
```

```{python}
# YOUR CODE HERE
```
### Data Collection & Cleaning

We already scraped the data from the EIA website, so we can use that data directly. We just need to read it in from the csvs it has been saved in.

```{python}
# YOUR CODE HERE
#Find all csv files in the directory, read them in, and store them in a single df
energy_df = pd.read_csv("./datasets/cleaned/cleaned_combined_data_with_derived_fixed.csv")
energy_df.head()
```

### Data Organization & Conditioning

Here I just check number of rows and columns.

```{python}
# YOUR CODE HERE
# Number of rows
print("Rows: ", energy_df.shape[0])
# Number of columns
print("Columns: ", energy_df.shape[1])
```


Columns Types
```{python} 
#Data types
energy_df.dtypes
```


### Data Storage

Find the number of missing values in each column.
```{python}
# YOUR CODE HERE
energy_df.isna().sum()
```


Shows the fuel types that have missing cost values
```{python}
# YOUR CODE HERE
missing_cost = energy_df[energy_df["cost"].isna()]
missing_cost["fuelTypeDescription"].unique()
```


Finds the counts for each fuel type
```{python}
# YOUR CODE HERE
missing_cost["fuelTypeDescription"].value_counts()
```


## Exploratory Data Analysis (EDA)

- What are the high-level patterns?
- What intuitive insights emerge?
- Is the data consistent across sources?
- What assumptions are we making?
- What new questions arise from the EDA?

 **Note**: Be mindful of *confirmation bias*—avoid interpreting data only to support your initial hypothesis.
```{r}
# YOUR CODE HERE
```

```{python}
# YOUR CODE HERE

```
## Model Planning

- **Model Type**: Classification, clustering, regression—what fits best?
- **Success/Failure Definitions**: Refine what constitutes a good or bad outcome.
- **EDA for Modeling**:
  - Identify relevant variables
  - Explore correlations
  - Apply domain knowledge
  - Verify assumptions
  - Prototype modeling ideas
```{r}
# YOUR CODE HERE
```

```{python}
# YOUR CODE HERE
```
## Model Building

- What models are being trained?
- How are they tested?
- What metrics are used to evaluate performance?
```{r}
# YOUR CODE HERE
```

```{python}
# YOUR CODE HERE
```
## Communicating Results

- Are the results **robust** across different scenarios?
- Are they **statistically significant**?
- Why do these results **matter** to stakeholders?
- How do they compare to your definitions of success and failure?

## Operationalization

- **Presentation**:
  - Who is the audience?
  - Is the code and analysis well-documented and replicable?
- **Deployment**:
  - Are models deployed on live data?
  - Are they behaving as expected?

MCDA and MCDA stats as graphs
```{r}
# Store criteria weights
critWeight <- tibble(
  criterion = c("Cost ($/MWh)", "CO2_Intensity_kg_per_MWh", "efficiency%", "CtoE_Ratio"),
  direction = c("cost","cost","benefit","cost"),
  weight = as.integer(c(35,35,15,15))
)

# Build tibble for pie chart display
myPie <- critWeight %>%
  mutate(label = c("Cost", "CO2 Intensity", "Thermal Efficiency", "CtoE Ratio"))

# Pie chart for criteria weights
ggplot(myPie, aes(x="", y=weight, fill=criterion))+geom_col(width=1)+coord_polar(theta="y")+geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 5)+theme_void()+theme(legend.position = "none")
```

```{r}
# select only intended criteria, pivot on criteria, set values to be their averages, then merge weights
enerlong <- py$energy_df %>%
  select(fueltypeid, `Cost ($/MWh)`, `efficiency%`, CO2_Intensity_kg_per_MWh, CtoE_Ratio) %>%
  pivot_longer(-fueltypeid, names_to = "criterion", values_to = "value") %>%
  mutate(
    fueltype = case_when(
      fueltypeid == "WNT" ~ "Wind",
      fueltypeid == "NUC" ~ "Nuclear",
      fueltypeid == "NG" ~ "Natural Gas",
      fueltypeid == "TSN" ~ "Solar",
      fueltypeid == "COW" ~ "Coal",
      fueltypeid == "PET" ~ "Petrol",
    )
  ) %>%
  group_by(criterion, fueltype) %>%
  group_modify(~{tibble(value = mean(.x$value, na.rm = TRUE))}) %>%
  ungroup %>%
  left_join(critWeight, by = "criterion")

# handle infinites and find normals
norm <- enerlong %>%
  group_by(criterion, direction) %>%
  mutate(
    value = ifelse(is.infinite(value), 1, value),
  ) %>%
  mutate(
    rng = max(value, na.rm = TRUE) - min(value, na.rm = TRUE),
    norm = case_when(
      is.na(value) ~ case_when(direction == "cost" ~ 0, direction == "benefit" ~ 1),
      rng == 0 ~ 1, 
      TRUE ~ (value - min(value, na.rm = TRUE)) / rng,
    )
  ) %>%
  ungroup()

# find weighted scores 
weighted <- norm %>%
  group_by(criterion, direction)%>%
  mutate(
    norm = ifelse(direction == "cost", 1 - norm, norm)
  ) %>%
  ungroup() %>%
  mutate(weighted = (norm * weight) / 100) 

# find weighted sums
wsm <- weighted %>%
  group_by(fueltype) %>%
  summarise(WSM_score = sum(weighted), .groups = "drop") %>%
  arrange(desc(WSM_score))
wsm

# Bar chart showing Normalized Performance Values
ggplot(norm, aes(x=fueltype, y=norm, fill = criterion))+geom_col(position = position_dodge(width = 0.8), width = 0.7)+labs(title = "Normalized Performance Values", x = "Fuel Type", y = "Normalized Average Values")+scale_fill_discrete(name = "Performance Value")

# Bar chart showing Weighted Performance Values
ggplot(weighted, aes(x=fueltype, y=weighted, fill = criterion))+geom_col(position = position_dodge(width = 0.8), width = 0.7)+labs(title = "Weighted Performance Values", x = "Fuel Type", y = "Weighted Values")+scale_fill_discrete(name = "Performance Value")

# bar chart showing WSM scores in order
ggplot(wsm, aes(x=reorder(fueltype, -WSM_score), y=WSM_score, fill = fueltype))+geom_col()+labs(title = "Final Rankings", x = "Fuel Type", y = "WSM Score")+ylim(0, 1)+theme(legend.position = "none")
print(norm, n=24)
```