---
title: "Efficient green energy sources report"
author: "Alejandro Barajas, Add your names here"
date: today
format: pdf
editor: source
---

# R and Python libary set up

```{r}
#| label: r setup
#| include: false
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, results = "hold")
library(tidyverse)
library(reticulate)
```

```{python}
#| label: python setup
#| include: false
import pandas as pd
```

- **Overall goal**
The goal is to identify which sources deliver the most energy per dollar while minimizing environmental impact

## Data Preparation
```{r}
# YOUR CODE HERE
```

```{python}
# YOUR CODE HERE
```
### Data Collection & Cleaning
- What sources are used?
- What cleaning steps are necessary?
```{r}
# YOUR CODE HERE
```

```{python}
# YOUR CODE HERE
```
### Data Organization & Conditioning
- How is the data structured?
- Are there transformations or feature engineering steps?
```{r}
# YOUR CODE HERE
```

```{python}
# YOUR CODE HERE
```
### Data Storage
- Where and how is the data stored?
- Is it accessible and secure?
```{r}
# YOUR CODE HERE
```

```{python}
# YOUR CODE HERE
```
## Exploratory Data Analysis (EDA)

- What are the high-level patterns?
- What intuitive insights emerge?
- Is the data consistent across sources?
- What assumptions are we making?
- What new questions arise from the EDA?

 **Note**: Be mindful of *confirmation bias*—avoid interpreting data only to support your initial hypothesis.
```{r}
# YOUR CODE HERE
```

```{python}
# YOUR CODE HERE
```
## Model Planning

- **Model Type**: Classification, clustering, regression—what fits best?
- **Success/Failure Definitions**: Refine what constitutes a good or bad outcome.
- **EDA for Modeling**:
  - Identify relevant variables
  - Explore correlations
  - Apply domain knowledge
  - Verify assumptions
  - Prototype modeling ideas
```{r}
# YOUR CODE HERE
```

```{python}
# YOUR CODE HERE
```
## Model Building

- What models are being trained?
- How are they tested?
- What metrics are used to evaluate performance?
```{r}
# YOUR CODE HERE
```

```{python}
# YOUR CODE HERE
```
## Communicating Results

- Are the results **robust** across different scenarios?
- Are they **statistically significant**?
- Why do these results **matter** to stakeholders?
- How do they compare to your definitions of success and failure?

## Operationalization

- **Presentation**:
  - Who is the audience?
  - Is the code and analysis well-documented and replicable?
- **Deployment**:
  - Are models deployed on live data?
  - Are they behaving as expected?